Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-24 paper/current/main.tex
edit figure 2 caption
cover comment on page 5

%
% Figure - Gravity law performance
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/GravityLawFig/gravity_performance.pdf}
	\caption{
		\textbf{Embedding distance outperforms geographic distance.}
		\textbf{a.} embedding distance (cosine distance between organization vectors) better explains the expected flux between organization pairs than geographic distance (\textbf{b}).
		The red line is the line of the best fit.
		Black dots are mean flux across binned distances.
		Error bars are 99\% confidence intervals.
		The lightness of each hex bin indicates the frequency of organization pairs within it.
		\textbf{c-d.} Predictions of flux between organization pairs made using embedding distance outperform those made using geographic distance.
		Boxplots show distribution of actual flux for binned values of predicted flux.
		Box color corresponds to the degree to which the distribution overlaps $y = x$.
		``RMSE'' is root-mean-squared error between the actual and predicted values.
		\textbf{e-f.} Predictions made using model parameters learned from a subset of domestic (within-country) organization pairs.
		Embedding distance outperforms those made for geographic distance.
		\textbf{g-h.} Embedding distance also outperforms geographic distance when predicting flux using model parameters estimated from only international mobility (across-country organization pairs).
		Predictions are not meaningfully different if estimates are learned on all pairs of organizations (Fig. \ref{fig:supp:gravity_prediction_subsets}).
		These predictions are robust whether the model is fit on the global data, or on the domestic (Fig.~\ref{fig:gravity_performance}e-h) or international (Fig.~\ref{fig:supp:gravity_prediction_subsets}) subsets.
	}
	\label{fig:gravity_performance}
\end{figure}

@@

The embedding space also allows us to \textit{zoom in} on subsets and re-project them in order to reveal more local relationships. 
For example, re-projecting organizations located in Western, Southern, and Southeastern Asia (Fig.~\ref{fig:projection}b) reveals a gradient of countries between Egypt to Philippines that corresponds largely to geography, but with some peculiar exceptions. 
For example, countries like Iraq are nearer to some southeast-Asian countries like Malaysia, with a cosine distance between the nation's aggregate vectors of $d = 0.29$, than to nearer and more religiously-similar nations like Saudi Arabia ($d = 0.50$) and Qatar ($d = 0.51$). 
Linguistic and historical ties are also conspicuous.
For example, we observe that Spanish-speaking post-colonial Latin American nations are positioned near Spain (Fig.~\ref{fig:projection}c), rather than nearby Portuguese-speaking nation ($d = 0.34$ vs. $d = 0.52$ for Mexico and $d = 0.39$ vs. $d = 0.49$ for Chile) reflecting linguistic and cultural constraints.
Similarly, North-African countries that were once under French rule such as Morocco are nearer to France ($d =0.35$) than to similarly-distant European countries as Spain ($d = 0.40$) and Italy ($d = 0.53$). 
Comparable structural patterns exist even within a single country.
For example, at the global level (Fig~\ref{fig:projection}a), organizations within Quebec in Canada are located nearer the French cluster ($d = 0.65$) than the rest of the Canadian universities ($d = 0.88$).

>>>>

%
% Figure - Gravity law performance
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/GravityLawFig/gravity_performance.pdf}
	\caption{
		\textbf{Embedding distance outperforms geographic distance.}
		\textbf{a.} embedding distance (cosine distance between organization vectors) better explains the expected flux between organization pairs than geographic distance (\textbf{b}).
		The red line is the line of the best fit.
		Black dots are mean flux across binned distances.
		Error bars represent 99\% confidence intervals.
		The lightness of each hex bin indicates the frequency of organization pairs within it.
		\textbf{c-d.} Predictions of flux between organization pairs made using embedding distance outperform those made using geographic distance.
		Boxplots show distribution of actual flux for binned values of predicted flux.
		Box color corresponds to the degree to which the distribution overlaps $y = x$.
		``RMSE'' is root-mean-squared error between the actual and predicted values.
		\textbf{e-f.} Predictions made using model parameters learned from a subset of domestic (within-country) organization pairs.
		Embedding distance outperforms those made for geographic distance.
		\textbf{g-h.} Embedding distance also outperforms geographic distance when predicting flux using model parameters estimated from only international mobility (across-country organization pairs).
		Predictions are not meaningfully different if estimates are learned on all pairs of organizations (Fig. \ref{fig:supp:gravity_prediction_subsets}).
		These predictions are robust whether the model is fit on the global data, or on the domestic (Fig.~\ref{fig:gravity_performance}e-h) or international (Fig.~\ref{fig:supp:gravity_prediction_subsets}) subsets.
	}
	\label{fig:gravity_performance}
\end{figure}

@@

The embedding space also allows us to \textit{zoom in} on subsets and re-project them in order to reveal more local relationships. 
For example, re-projecting organizations located in Western, Southern, and Southeastern Asia (Fig.~\ref{fig:projection}b) reveals a gradient of countries between Egypt to Philippines that corresponds largely to geography, but with some peculiar exceptions based on cultural and religious spectrum.
For example, countries like Iraq are nearer to some southeast-Asian countries like Malaysia, with a cosine distance between the nation's aggregate vectors of $d = 0.29$, than to nearer and more religiously-similar nations like Saudi Arabia ($d = 0.50$) and Qatar ($d = 0.51$). 
Linguistic and historical ties are also conspicuous.
For example, we observe that Spanish-speaking post-colonial Latin American nations are positioned near Spain (Fig.~\ref{fig:projection}c), rather than nearby Portuguese-speaking nations ($d = 0.34$ vs. $d = 0.52$ for Mexico and $d = 0.39$ vs. $d = 0.49$ for Chile) reflecting linguistic and cultural constraints.
Similarly, North-African countries that were once under French rule such as Morocco are nearer to France ($d =0.35$) than to similarly-distant European countries as Spain ($d = 0.40$) and Italy ($d = 0.53$). 
Comparable structural patterns exist even within a single country.
For example, at the global level (Fig~\ref{fig:projection}a), organizations within Quebec in Canada are located nearer the French cluster ($d = 0.65$) than the rest of the Canadian universities ($d = 0.88$).
